---
title: "R Notebook"
output: html_notebook
---

HF data exploration


```{r message=FALSE, warning=FALSE, include=FALSE}
library(corrplot)
library(psych)
library(vioplot)
library(gplots)
library(readxl)
library(gridGraphics)
library(ggpubr)
library(caret)
library(leaps)
library(PerformanceAnalytics)
library(factoextra)
library(vcd)
library(DataExplorer)
```

The outcome variable is Time to Tx (hrs) and can be obtained from columns 154 and 155. The ID column is column 1.  Catigorical variables are from 2:133, numerical variables run from 134:153.
```{r}

setwd("~/Desktop/research/HF_Study/")
TimetoTreatment = NULL
TimetoTreatment <-read_excel("80percData_IncludingTimeToTx_04MAR20.xlsx",header = TRUE)
dim(TimetoTreatment)
#[1] 382  31
TTData = NULL
TTData = TimetoTreatment

#Outcome
names(TTData[,155])
Y_H = TTData[,155]
Y_HB = TTData[,154]
user = TTData[,1]

barplot(table(TTData[,154]), space = 0.8, col = "chartreuse4")
#gf1 = goodfit(Y_HB, "poisson")
#rootogram(gf1, xlab = "", rect_gp = gpar(fill = "chartreuse4"))
```


```{r}
plot_bar(TTData)
```


```{r}
multi.hist(TTData[,134:154]) 
multi.hist(log(TTData[,134:154]))

```
Looking for normality
```{r}
plot_qq(log(TTData[,134:154]))
```


```{r}
#more examples:
#http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram


#First we have to run the cor function on our data:
TT_cor_values <-cor(TTData[,134:154],method = "kendall")

#Here is an example with colors and circles
corrplot(TT_cor_values, method="circle")

#First we have to run the cor function on our data:
TT_cor_values <-cor(TTData[,134:154],method = "pearson")

#Here is an example with colors and circles
corrplot(TT_cor_values, method="circle")

chart.Correlation(TTData[,134:154],histogram=TRUE, col="grey10", pch=1, main="HF data")
```

Working on developing methods
```{r}

hf_status <- factor(TTData$hf_status)
Y_HB_T = as.numeric(unlist(Y_HB))
t_test_data = t.test(Y_HB_T~hf_status)
wilcox_test = wilcox.test(Y_HB_T~hf_status)
#t_test_data
plotmeans(Y_HB_T ~ hf_status, xlab = "Heart Failure Status",ylab = "Time to Treat", main = "Mean Plot Heart Failure Status vs Time to Treat \nwith 95% CI")

model <- lm(Y_HB_T ~ factor(TTData$hf_status), data = TTData)
anova(model)
summary(model)$coef




```
How does what we see translate into a scatter plot.
```{r}
plot(Y_HB_T,col = hf_status,ann=FALSE )
plot(log(Y_HB_T),col = hf_status,ann=FALSE )

#hf_status
```

abnp2_279,admhr,acreatin2_4d5,abun2_371

Trying to use a function to save plots
```{r}

plot_maens<-function(numerical_variable = NULL,catigorial_variable = NULL){
  numerical_variable <- as.numeric(unlist(numerical_variable))
  catigorial_variable <- factor(catigorial_variable)
  
  xtext = paste(names(catigorial_variable),"Status",sep=" ")
  ytext = names(numerical_variable)
  plotmeans(Y_HB_T ~ catigorial_variable, xlab = xtext,ylab = ytext, main = "Mean Plot Heart Failure Status vs Time to Treat \nwith 95% CI")
}

plot_maens(TTData$abnp2_279,TTData$hf_status)



```

Using a function to run statistics

```{r}

gnerating_p_values <-function(numerical_variable = NULL,catigorial_variable = NULL){
  
  catigorial_variable <- factor(catigorial_variable)
  numerical_variable = as.numeric(unlist(numerical_variable))
  t_test_data = t.test(numerical_variable~catigorial_variable)
  #t_test_data
  wilcox_test = wilcox.test(numerical_variable~catigorial_variable)
  model <- lm(numerical_variable ~ factor(catigorial_variable), data = TTData)
  return(wilcox_test)
  
}

gnerating_p_values(TTData$abnp2_279,TTData$hf_status)

```

Working do not delete!
```{r}

stuff = list(names(TTData))
p = NULL
plot_list = NULL
#master_plot_maker_cat <- function(){
plot_list = list()
j = 0
for (i in 1:131) {
    #j = j + 1
    p = ggline(TTData, x = stuff[[1]][i], y = stuff[[1]][154],add = c("mean_se"))
    n = p + stat_compare_means(label.y = 100,label.x = 1)
    plot_list[[i]] = n
}

pdf("Time_to_treat_plots.pdf")
par(mfrow=c(26,6))
plot_list
dev.off()

# wilcox_test = wilcox.test(Y_HB_T ~ TTData$panas___6)
# wilcox_test
# t.test(log(Y_HB_T) ~ TTData$panas___6)
# 
# wilcox_test = wilcox.test(Y_HB_T ~ TTData$pches2_e4a___1)
# 
# p = ggline(TTData, x = stuff[[1]][2], y = stuff[[1]][154],add = c("mean_se"))
# p + stat_compare_means(label.y = 120)



```

Doing the same for other sets
abnp2_279,admhr,acreatin2_4d5,abun2_371

```{r}
stuff = list(names(TTData))
p = NULL
plot_list = NULL
#master_plot_maker_cat <- function(){
plot_list = list()
j = 0
for (i in 1:131) {
    #j = j + 1
    p = ggline(TTData, x = stuff[[1]][i], y = stuff[[1]][145],add = c("mean_se"))
    n = p + stat_compare_means(label.y = 2,label.x = 1)
    plot_list[[i]] = n
}

pdf("acreatin2_4d5_plots.pdf")
par(mfrow=c(26,6))
plot_list
dev.off()
```

Here I am making an interaction plot that should allow me to look at more things
```{r}

percough2_1e2_3 = factor(TTData$percough2_1e2___3)
pches2_e4a_1 = factor(TTData$pches2_e4a___1)
interaction.plot(pches2_e4a_1,percough2_1e2_3, Y_HB_T, type = "b", col = c(1:3), leg.bty = "o", leg.bg = "beige", lwd = 2, pch = c(18,24,22), xlab = "pches2_e4a___1", ylab = "Time to Treat", main = "Interaction Plot")
```


```{r}

hf_status <- factor(TTData$hf_status)
Y_HB_T = as.numeric(unlist(Y_HB))
coplot(Y_HB_T ~ hf_status | pches2_e4a_1, data = TTData,panel = panel.smooth, rows = 1)

```

Trying to develop a linear model to try to predict time to treat
First I have to prepare the data
```{r}
#Y_HB = TTData[,154]
#TT_cor_values <-cor(TTData[,134:154],method = "kendall")
Rawdata = TTData[,134:154]
betas = NULL
rows = nrow(Rawdata)
betas = matrix(0,nrow = rows,ncol = ncol(TTData[,134:154]))

Y_HB = TTData[,154]
outcome = as.numeric(unlist(Y_HB))

for(i in 1:ncol(TTData[,134:154])){
  betas[,i] = as.numeric(unlist(Rawdata[,i]))
}

betas = as.data.frame(betas)
colnames(betas) = names(TTData[,134:154])
betas




```

Now lets build a general model
```{r}
lmTmp = NULL
#betas = cbind(betas,outcome)
lmTmp <- lm(betas$`Time to Tx (min)` ~ .,data = betas)
summary(lmTmp)

VImp_lm = varImp(lmTmp, scale = FALSE)
VImp_lm

errTmp <- sqrt(mean((betas$`Time to Tx (min)` - predict(lmTmp))^2))
errTmp

#not fruitful
lmTmp_lite <- lm(betas$`Time to Tx (min)` ~ betas$admhr + betas$admresp + betas$achlor2_945 + betas$abicarb2_60a,data = betas)
summary(lmTmp_lite)
lmTmp_abicarb2 = lm(betas$`Time to Tx (min)` ~ betas$abicarb2_60a,data = betas)
summary(lmTmp_abicarb2)

plot(y = betas$`Time to Tx (min)`, x = log(betas$abicarb2_60a))
abline(lmTmp_abicarb2,col = "red")
```

```{r Subset selction}


#*********************************
# Subset selction and ranking
#*********************************
#Running regsubsets on the dataset to get a sense for what methods contribute most
summaryMetrics <- NULL
whichAll <- list()

  for ( myMthd in c("exhaustive", "backward", "forward") ) {
    #lmTmp <- lm(betas$`Time to Tx (min)` ~ .,data = betas)
    rsRes <- regsubsets(betas$`Time to Tx (min)` ~.,betas,method=myMthd,nvmax=ncol(betas))
    summRes <- summary(rsRes)
    whichAll[[myMthd]] <- summRes$which
    for ( metricName in c("rsq","rss","adjr2","cp","bic") ) {
      summaryMetrics <- rbind(summaryMetrics,
        data.frame(method=myMthd,metric=metricName,
                  nvars=1:length(summRes[[metricName]]),
                  value=summRes[[metricName]]))
    }
  }

ggplot(summaryMetrics,aes(x=nvars,y=value,shape=method,colour=method)) + geom_path() + geom_point() + facet_wrap(~metric,scales="free",ncol=5) +   theme(legend.position="top")




old.par <- par(mfrow=c(1,3),ps=16,mar=c(5,12,5,1))

  for ( myMthd in names(whichAll) ) {
    image(1:nrow(whichAll[[myMthd]]),
          1:ncol(whichAll[[myMthd]]),
          whichAll[[myMthd]],xlab="N(vars)",ylab="",
          xaxt="n",yaxt="n",breaks=c(-0.5,0.5,1.5),
          col=c("white","gray"),main=paste(myMthd))
    axis(1,1:nrow(whichAll[[myMthd]]),rownames(whichAll[[myMthd]]))
    axis(2,1:ncol(whichAll[[myMthd]]),colnames(whichAll[[myMthd]]),las=2)
  }
par(old.par)

#*********************************
# Subset selction and ranking
#*********************************



```

```{r}

#TTData[,134:154]
#here we are computing PCA
prcompTmp <- prcomp(TTData[,134:154],center = TRUE,scale. = TRUE)

#Here we have a plot of the variance by principle componant
plot(prcompTmp)

fviz_eig(prcompTmp, addlabels=TRUE, ylim=c(0,60), geom = c("bar", "line"), barfill = "gold", barcolor="grey",linecolor = "red", ncp=10)+
labs(title = "Cancer All Variances - PCA",
         x = "Principal Components", y = "% of variances")

fviz_pca_biplot(prcompTmp, col.ind = TTData$`90 min cut-off`, col="black",
                palette = "jco", geom = "point", repel=TRUE,
                legend.title="0 = Benign or 1 = Malignant", addEllipses = TRUE)
```
